Supplementary Information

### 2021-08-06 doc/manuscript/response_to_reviews.Rmd

 L222-224:
Dealing with tree and forest mortality is difficult. However, climate change is predicted to affect mortality in many forests. In turn, mortality can impact the growth of the remaining trees depending on the stature and location of the dead tree. This point should be addressed- the article presents the relationship between growth and climate only for the surviving trees. How does excluding growth data from the analysis potentially bias the results?

*increased mortality--> decreased competition --> increased growth. This should be mentioned in the discussion.*

P19-20 (especially L301-303):
I think it could be made more clear that detrending was used only in the process of selecting strongly related climate variables, and not in the GLS models. Additionally, a comment on how the variability in the application of detrending method used here (thin plate splines) would impact climate variable selection would be warranted. There is a persistent question in dendrochronology about the amount of climate signal gets removed through the process of detrending.

**In the GLS models, we use a temporal autocorrelation structure on raw data. We have revised the GLS model description to make this more clear: "We next combined the primary climate drivers in temperature and precipitation variable groups (included in all models) and DBH (included in models with DBH and DBH-climate interactions) into linear mixed-effects models [function *lme* in the R package *nlme*, @pinheiro_nlme_2021], with core identity as a random intercept. We included year as a continuous time covariate for the within-group correlation structure (function *corCAR1*) to account for temporal autocorrelation (similar to how detrending would). We will refer to this model as a generalized least squares (GLS) model (Fig. 1).Within the GLS models, our response variables were raw, log-transformed growth estimates (as opposed to residuals): log[RW], log[BAI], or log[$\Delta$ AGB].**

### 2021-08-06 doc/manuscript/ForestGEO_treerings.Rmd
small change to model description

We next combined the primary climate drivers in temperature and precipitation variable groups (included in all models) and DBH (included in models with DBH and DBH-climate interactions) into linear mixed-effects models [function *lme* in the R package *nlme*, @pinheiro_nlme_2021], with core identity as a random intercept and year as a continuous time covariate for the within-group correlation structure (function *corCAR1*). 
We will refer to this model as a generalized least squares (GLS) model (Fig. `r schematic`).
Within the GLS models, our response variables were raw, log-transformed growth estimates (as opposed to residuals): log[RW], log[BAI], or log[$\Delta$ AGB].
Prior to running the models, we checked for collinearity among the candidate variables using the *vifstep* function [@naimi_where_2014]. Our analysis code was programmed to remove any variable with a variance inflation factor > 3, but none required removal.

>>>>

We next combined the primary climate drivers in temperature and precipitation variable groups (included in all models) and DBH (included in models with DBH and DBH-climate interactions) into linear mixed-effects models [function *lme* in the R package *nlme*, @pinheiro_nlme_2021], with core identity as a random intercept. We included year as a continuous time covariate for the within-group correlation structure (function *corCAR1*) to account for temporal autocorrelation (similar to how detrending would). 
We will refer to this model as a generalized least squares (GLS) model (Fig. `r schematic`).
Within the GLS models, our response variables were raw, log-transformed growth estimates (as opposed to residuals): log[RW], log[BAI], or log[$\Delta$ AGB].
Prior to running the models, we checked for collinearity among the candidate variables using the *vifstep* function [@naimi_where_2014]. Our analysis code was programmed to remove any variable with a variance inflation factor > 3, but none required removal.
